\begin{abstract}
Large Language Models (LLMs) can read an OpenAPI specification and synthesise REST API
tests automatically, and a rapidly growing literature reports that they outperform
search-based tools on live-API error counts. Across \textbf{59} surveyed papers
(2018--2026) we find, however, that three questions remain unanswered: no study compares
an LLM generator against \emph{both} a human-style manual suite \emph{and} a mature tool
on the same APIs; none computes fault-detection \emph{Recall} against a known,
pre-seeded-fault ground truth (fault counts are reported as raw 500-error tallies on live
systems); and none reports a per-endpoint edge-case metric.

We address all three with a controlled, blind comparison of three arms --- an LLM
(Claude Sonnet 4.6), an independently authored manual EP/BVA suite, and EvoMaster 6.0.0 ---
on three EvoMaster Benchmark (EMB) REST APIs comprising \RQIopsTotal{} operations, into which
\RQIIn{} faults are seeded by mutation so that the denominator of Recall is known exactly.
Our results are pointedly mixed, and the mixture is the contribution. The LLM reaches
\RQIcoveragePct\% endpoint coverage ($\RQIopsCovered/\RQIopsTotal$, $p=\RQIpvalue$) and
produces \RQIIItotalLlm{} edge-case scenarios against manual's \RQIIItotalManual{}
($p=\RQIIIpvalue$, rank-biserial $=\RQIIIrankBiserial$) --- it is decisively more
\emph{prolific}. On the ground truth the ordering is EvoMaster \RQIIrecallEvo{} $>$ LLM
\RQIIrecallLlm{} $>$ manual \RQIIrecallManual{}: the LLM significantly out-detects the
independent human-style suite (McNemar $p=\RQIImcnemarLMp$, its kills a strict superset) yet
is itself out-detected by a search-based tool from 2019 ($p=\RQIImcnemarLEp$; both hold at
$\alpha=\alphaRaw$ but not under Bonferroni, so we read them as directional).

We trace the ordering to two composing mechanisms the coverage-centric literature cannot
observe. First, oracle \emph{kind}: the LLM's and the manual suite's spec-derived
status-code oracles accept any response in a plausible class, so they are structurally blind
to computational faults that alter a returned value while preserving its HTTP status ---
on the arithmetic-heavy subject both score zero while EvoMaster's recorded-value regression
oracles catch twelve. Second, within the status-oracle regime, \emph{volume}: where faults
are status-visible, the LLM's larger suite catches more of them than the leaner independent
manual suite (its kills a strict subset). Test volume and fault-finding power are thus
distinct, and per-endpoint scenario counts --- the metric the field is trending toward ---
actively mislead when reported without a ground-truth Recall alongside them.
\end{abstract}

\begin{IEEEkeywords}
REST API testing, large language models, mutation testing, test oracles, EvoMaster,
empirical software engineering
\end{IEEEkeywords}
