\section{Related Work}
\label{sec:related}

Our evidence base is a five-member systematic review merged into 59 unique papers
(2018--2026) after de-duplication, screened against pre-registered inclusion and exclusion
criteria with a PRISMA flow recorded per member. Table~\ref{tab:related} lists the ten
representative studies that define the state of the art.

\begin{table}[tb]
\centering
\caption{Representative studies from the 59-paper merge. Every fault figure in the
right-hand column is a count on a live system; none is a Recall against a known fault
population.}
\label{tab:related}
\footnotesize
\setlength{\tabcolsep}{3pt}
\begin{tabular}{@{}p{2.8cm}p{1.6cm}p{1.4cm}p{1.7cm}p{3.85cm}@{}}
\toprule
\textbf{Study (year)} & \textbf{Generator} & \textbf{Dataset} & \textbf{Metric} & \textbf{Headline result} \\
\midrule
RESTGPT~\cite{restgpt} (2024)      & GPT-3.5              & 9 real APIs   & rule P/R/F1        & 97\% rule precision; 72.7\% valid inputs \\
KAT~\cite{kat} (2024)              & GPT-3.5-turbo        & 12 services   & status-code cov.\  & 74.9\% (+15.7\,pp) \\
RESTSpecIT~\cite{restspecit} (2024)& DeepSeek / GPT       & 10 PRAB APIs  & route disc., 5xx   & 88.6\% routes; 5xx in 4 APIs \\
AutoRestTest~\cite{autoresttest} (2025) & GPT-3.5 + MARL  & 12 services   & code cov., 500s    & 58.3\% line; 42 500s vs EvoMaster 20 \\
LlamaRestTest~\cite{llamaresttest} (2025) & Llama3-8B (ft) & 12 services  & code cov., faults  & 55.8\% method; 204 faults vs 130 \\
LogiAgent~\cite{logiagent} (2025)  & GPT-4o-mini          & 12 systems    & cov., logical bugs & 71.78\% line; 49 crashes \\
RESTifAI~\cite{restifai} (2026)    & GPT-4.1-mini         & 7 services    & operation cov.\    & 128/134 ops $\approx$ 95.5\% \\
EvoMaster~\cite{evomaster} (2019)  & search-based         & 3 services    & stmt.\ cov., \textbf{vs manual} & 38 bugs; 41\% generated vs \textbf{82\% manual} \\
Morest~\cite{morest} (2022)        & model-based          & 6 projects    & cov., bugs         & 44 bugs (13 new) \\
No-Time-to-Rest~\cite{notimetorest} (2022) & 10 tools     & 20 EMB svcs   & line/branch, 500s  & EvoMaster-WB 52.76\% line \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Four Patterns}

\textbf{1. GPT dominates the LLM line.}
Nearly every LLM study uses an OpenAI
model~\cite{kat,restspecit,restgpt,autoresttest,logiagent,restifai,apitestgenie}, and
open-source models appear only sporadically~\cite{llamaresttest}. Our use of Claude Sonnet 4.6
therefore adds a data point outside the family that supplies almost all current evidence,
which is a partial contribution to GAP-T.

\textbf{2. Metrics are coverage plus error counts.}
Semantic and edge-case quality is rarely isolated. Only KAT~\cite{kat} splits 2xx from 4xx,
and only RESTSpecIT~\cite{restspecit} reports 5xx. No study reports edge cases per endpoint,
so there is no baseline against which a suite's negative-testing breadth can be judged.

\textbf{3. Baselines are automated-only.}
LLM tools are compared against EvoMaster or Morest and reported to beat them on server
errors~\cite{llamaresttest,autoresttest}. No LLM study includes a manual suite. The single
manual comparison in the field remains EvoMaster's own~\cite{evomaster}, which found the tool
well below manual (41\% vs 82\%). No subsequent LLM paper has revisited that comparison, even
as the field's claims moved toward human parity.

\textbf{4. Fault detection has no denominator.}
Every fault figure in Table~\ref{tab:related} is a count on a live
API~\cite{llamaresttest,autoresttest,logiagent,notimetorest,evomaster,morest}. We regard this
as the most consequential methodological problem in the area, since error counts and Recall
can be ordered differently: a suite that hammers one endpoint hard may log many 500s while
missing most of the fault population. Mutation testing is standard elsewhere in software
engineering~\cite{pit,offutt}, and its absence here leaves the comparative claims untestable.

\subsection{Positioning}
Table~\ref{tab:gapmap} maps each gap to the evidence establishing it. Our design targets
GAP-C and GAP-D jointly, since adding both a manual arm and a known fault population is what
makes the three-way comparison meaningful, and treats GAP-M as the metric facet through which
the result is interpreted.

\begin{table}[tb]
\centering
\caption{Gap map: what the 59-paper base does and does not establish.}
\label{tab:gapmap}
\footnotesize
\begin{tabular}{@{}p{1.5cm}p{7.6cm}p{2.0cm}@{}}
\toprule
\textbf{Gap} & \textbf{Evidence} & \textbf{Status} \\
\midrule
GAP-C & \cite{evomaster} is the only manual comparison (non-LLM, scored below manual); \cite{llamaresttest,autoresttest,notimetorest} automated-only & Confirmed \\
GAP-D & \cite{llamaresttest,autoresttest,logiagent,notimetorest,evomaster,morest} report counts, never Recall & Confirmed \\
GAP-M & \cite{kat} (2xx/4xx) and \cite{restspecit} (5xx) partial; none per-endpoint & Confirmed \\
GAP-T & open-source / non-GPT models sporadic & Deferred (ablation) \\
\bottomrule
\end{tabular}
\end{table}
