\section{Method}
\label{sec:method}

The research questions, metrics, thresholds and statistical tests below were fixed in a
proposal frozen at instructor approval in Week~5, before any full-run data were seen. RQ2's
result contradicts the hypothesis we stated there.

\subsection{Research Questions and Hypotheses}

\textbf{RQ1 (absolute).} Do LLM-generated tests reach endpoint coverage $\geq 90\%$, and
which endpoint types are most often missed?
$H_0: \mu \leq 0.90$; $H_1: \mu > 0.90$. The 90\% threshold is bracketed by the literature,
which spans 88.6\% route discovery~\cite{restspecit} at the low end and 95.5\% operation
coverage~\cite{restifai} at the high end.
\emph{Test:} one-sample Wilcoxon signed-rank against 0.90, one-tailed.

\textbf{RQ2 (comparative).} How many pre-seeded faults do LLM tests detect, versus manual
and versus EvoMaster? $H_0$: $\text{rate}_{\text{LLM}} \leq
\max(\text{rate}_{\text{manual}}, \text{rate}_{\text{EvoMaster}})$; $H_1$: greater than
both. \emph{Test:} Friedman across arms, post-hoc pairwise Wilcoxon with Holm correction,
and a pooled per-fault McNemar test; effect size Cliff's $\delta$.

\textbf{RQ3 (comparative).} Do LLM tests produce more 4xx/5xx and boundary scenarios per
endpoint than manual design? $H_0$: $\text{median}_{\text{LLM}} \leq
\text{median}_{\text{manual}}$. \emph{Test:} paired Wilcoxon signed-rank across endpoints,
one-tailed; effect size rank-biserial.

We use $\alpha=\alphaRaw$ per RQ as pre-registered, and additionally report a Bonferroni
threshold across the three RQ families ($\alpha_{\text{adj}} \approx \alphaBonf$) as a
stricter second reading.

\subsection{Subjects}

We use three REST services from the EvoMaster Benchmark (EMB, LGPL-3.0), the shared corpus of
the field~\cite{notimetorest,autoresttest}: \texttt{rest-ncs} (\RQINcsTotal{} operations),
\texttt{rest-scs} (\RQIScsTotal{}), and \texttt{features-service} (\RQIFeaturesTotal{}),
totalling \RQIopsTotal{} operations. Each is deployed locally as a standalone JDK~8 Spring
Boot fat jar. The harness and EvoMaster run on JDK~17 and communicate with the subjects over
HTTP only. Provenance, license and operation counts are recorded in
\texttt{data/raw/README.md}.

\subsection{Fault Seeding and Ground Truth}
\label{sec:method-faults}

Faults are seeded by mutation, which supplies the denominator the literature lacks. We apply
standard PIT/Offutt operator families~\cite{pit,offutt}, namely relational replacement,
arithmetic replacement and negate/boundary conditionals, to the controller and core-logic
classes of each subject, where a fault is observable through the HTTP contract. Each mutation
is a single-token change, applied in isolation and recompiled into its own mutant jar. Mutants
that fail to compile are discarded, and the pristine source is restored before the next.

This yields \RQIIn{} compilable mutants: 70 in \texttt{rest-ncs}, 59 in \texttt{rest-scs}
and 4 in \texttt{features-service}. Each is catalogued with file, line and operator in
\path{data/faults/<sut>/catalog.json} and exported to \path{data/full_ground_truth.csv}.
The ground truth is derived from code, not from human annotation, so no inter-annotator
agreement statistic applies.

The \texttt{features-service} imbalance (4 mutants against 70 and 59) is a real property of
that service's arithmetic-light controller code rather than a sampling choice. We report it as
a generalisation limit instead of repairing it after the fact.

\subsection{The Three Arms}

All three suites are authored \emph{before} faults are seeded and are blind to the fault
catalog, so no arm can target a known mutant.

\textbf{LLM.} Claude Sonnet 4.6 (\texttt{claude-sonnet-4-6}), temperature $0$, top\_p $1$,
max\_tokens $4096$, fixed seed. Input is the OpenAPI specification only, so the arm is
black-box with no access to source. The prompt is frozen verbatim in
\texttt{scripts/llm/prompt\_template.md}. Per-subject transcripts and the tool-use evidence
establishing spec-only input are in \texttt{llm/transcripts/}. Output is REST-assured JUnit~5
tests.

\textbf{Manual.} An equivalence-partitioning and boundary-value-analysis suite designed from
the same specification and blind to faults, with the design procedure recorded in
\path{manual/methodology.md}. This suite
was authored by a separate, isolated model session rather than by a human cohort. We treat
that substitution as a construct-validity threat and discuss it in
Section~\ref{sec:threats}; throughout this paper the arm is labelled ``independently authored''
rather than ``human''.

\textbf{EvoMaster 6.0.0}, black-box mode, seed 42, three repeats:

\begin{center}
\texttt{-{}-blackBox true -{}-bbSwaggerUrl <spec>}\\
\texttt{-{}-outputFormat JAVA\_JUNIT\_5 -{}-maxTime 120s}
\end{center}

\subsection{Measurement}

\textbf{Coverage} is the number of operations exercised by at least one test, divided by the
total number of operations, and is computed from the specification together with the executed
request log.

\textbf{Fault Recall} is $\text{killed}/\RQIIn{}$. An arm kills a mutant if at least one test
that \emph{passes} on the original subject \emph{fails} on the mutant, so that the suite
observably distinguishes the faulty build from the correct one. Tests already failing on the
original are excluded as invalid oracles for that arm, which is why we record
\texttt{n\_oracle} per arm alongside each kill decision.

\textbf{Edge-case count} is the number of distinct negative, boundary and error-code scenarios
per endpoint, parsed from \texttt{// SCENARIO type=} tags in the suite source.

\textbf{Error-surface baseline.} A static analyser (\texttt{scripts/error\_surface.py})
enumerates, per endpoint, every source location that can emit a non-2xx response, classified
as \emph{declared} (explicit \texttt{status(4xx/5xx)}), \emph{framework} (typed
\texttt{@PathVariable} giving a Spring auto-400), or \emph{potential} (uncaught exception
giving a 500). Section~\ref{sec:intro} describes what this baseline is for.

\textbf{Error-behaviour key.} Separately from the static baseline, we record which distinct
(operation, status) behaviours each arm actually elicited at run time
(\texttt{scripts/error\_profile.py}). The denominator of \PDerrKey{} behaviours is the
\emph{union} of what the three arms observed, not an independently derived total. An arm that
scores \PDerrKey{}$/$\PDerrKey{} has therefore observed a superset of the others' behaviours;
it has not been shown to cover the full error surface. Section~\ref{sec:threats} states the
consequence for how this metric may be read.

\textbf{Produced versus detected.} Every request each suite issues is replayed through a
logging proxy against the live subject and its real HTTP status recorded
(\texttt{results/raw/traffic.csv}). We then report separately the number of tests and
scenarios \textbf{produced} and the number of faults and error behaviours actually
\textbf{detected} (\texttt{results/produced-vs-detected.json}). The separation matters because
Section~\ref{sec:results} shows the two orderings disagree, and conflating them is how a suite
that finds fewer faults can be reported as the better one.

\subsection{Reproducibility}
\label{sec:method-repro}

The pipeline is a chain in which each stage writes a committed artefact:

\begin{center}
\texttt{mutate.py} $\rightarrow$ \texttt{faults/*/catalog.json} $\rightarrow$
\texttt{run\_mutation.py} $\rightarrow$ \texttt{results/raw/*.csv} $\rightarrow$
\texttt{analyze.py} $\rightarrow$ \texttt{results/stats/summary.json} $\rightarrow$
\texttt{gen\_paper\_macros.py} $\rightarrow$ this paper
\end{center}

Every computed statistic in this document is a LaTeX macro bound by
\texttt{gen\_paper\_macros.py} to a field of \texttt{summary.json}, so re-running the
experiment moves the paper's text. A small number of fixed integers are written literally,
namely the per-subject mutant counts and figures quoted from the Week-7 pilot;
\texttt{scripts/check\_paper.py} fails the build if any \emph{derived} statistic is
hardcoded into a results section. We verified that \texttt{analyze.py} regenerates
\texttt{summary.json} byte-identically from the committed raw data, and that the analysis
notebook survives Restart~\&~Run~All with zero errors. Statistics use \texttt{scipy} and
\texttt{statsmodels}, with the pinned toolchain recorded in \texttt{env/tools.md}.
