\section{Discussion}
\label{sec:discussion}

\subsection{Why Volume and Effectiveness Came Apart}

Section~\ref{sec:results} reports a dissociation. The LLM produced
$\PDvolumeRatioLlmEvo\times$ EvoMaster's test volume, covered every operation, generated
\RQIIItotalLlm{} edge-case scenarios against the manual suite's \RQIIItotalManual{}, and still
killed fewer than half as many faults as EvoMaster. The explanation lies in the kind of
assertion each arm writes.

The LLM reads an OpenAPI specification, and a specification documents status codes: this path
returns 200, that one returns 400 on bad input. Tests synthesised from it therefore assert on
the status code and, at best, on the shape of the response body. The resulting oracle says
only that the response falls in the plausible class for the request. Our manual EP/BVA suite,
drawn from the same specification, writes assertions of the same kind. Both arms are
specification-derived status oracles.

Consider what our mutants do to such an oracle. A relational or arithmetic operator
replacement inside a numerical routine, such as \texttt{<} becoming \texttt{<=} or \texttt{/}
becoming \texttt{*}, changes a computed value. The endpoint still returns 200 with a wrong
number in the body, so a status oracle passes the mutant and the fault survives.

EvoMaster writes a different kind of oracle. Its black-box mode records the actual response of
the original system and emits assertions pinning those recorded values. This is a regression
oracle, and it fails whenever any observed value changes, whether or not the status changes
with it. Against value-changing mutants it is the right instrument.

The arms were therefore never measuring the same thing. The LLM and manual arms answer whether
the API responds in the documented class. EvoMaster answers whether it still computes what it
computed before. Our fault population consists almost entirely of value faults, so it bears on
the second question. EvoMaster's advantage should be read as evidence that a regression oracle
beats a status oracle on value faults, not as evidence that search-based generation is more
capable than an LLM.

The per-subject pattern is what this account predicts. If oracle kind drives the
result, EvoMaster's advantage should concentrate where faults are most value-like, and it
does. EvoMaster leads on \texttt{rest-ncs} (\RQIINcsEvo{} against \RQIINcsLlm{}), whose
controllers are numerical routines in which nearly every mutant shifts a computed value while
preserving the status. On \texttt{rest-scs}, which has more string and validation logic and so
more mutants that flip a response class, the LLM leads (\RQIIScsLlm{} against \RQIIScsEvo{}).

The \texttt{rest-ncs} kills make the mechanism concrete. All \NcsEvoKills{} mutants EvoMaster
kills there lie in a single file, \texttt{\NcsEvoKillFile.java}, an implementation of the
Bessel function. \NcsEvoKillsArith{} of them swap one arithmetic operator for another
(\texttt{/} for \texttt{*}, \texttt{-} for \texttt{+}) and the remaining \NcsEvoKillsRel{}
alter a numeric comparison. Every one leaves the request valid and the response status at 200,
changing only the number returned. That is the exact fault class a status oracle cannot
express an assertion about, and it is why both spec-derived arms score \RQIINcsLlm{} on this
subject.

Suite size does not explain the gap, and in fact runs against it. On \texttt{rest-ncs} the LLM
contributes \NcsOracleLlm{} tests that pass on the unmutated subject and therefore qualify as
oracles, and the manual suite \NcsOracleManual{}, against EvoMaster's \NcsOracleEvo{}.
EvoMaster kills \NcsEvoKills{} mutants with the smallest valid-oracle suite of the three while
the two larger suites kill none. Within this subject the relationship between test volume and
fault detection is not weak but inverted, which is what a ceiling set by oracle kind looks
like from the outside.

The LLM also does better where its oracle is adequate. Where being wrong means returning a
different status, it elicits the widest range of behaviours of any arm
(\PDerrTriggeredLlm{} against manual's \PDerrTriggeredManual{} and EvoMaster's
\PDerrTriggeredEvo{}), and it produces a genuine 500 crash. That key is arm-derived, so it
cannot tell us how much of the error surface any arm reached
(Section~\ref{sec:threats}); it does show that the LLM's greater volume converts into more
observed behaviour, but only among faults its oracle could see in the first place.

That gives the two-level picture: oracle kind sets a ceiling and volume distributes what falls
under it. The pilot behind
this study reported the LLM and manual arms killing identical mutant sets, but its manual
suite shared an author with the LLM suite, so that identity could not be interpreted.
Re-authoring the manual suite independently separates two effects the tie had conflated. On
\texttt{rest-ncs}, where every mutant is a value fault, the independent manual suite still
scores \RQIINcsManual{}, exactly like the LLM: no status-oracle suite, however authored, can
see a wrong number behind a 200. On \texttt{rest-scs}, where some mutants flip a response
class and are therefore status-visible, the suites separate sharply. The LLM kills
\RQIIScsLlm{} to the independent manual's \RQIIScsManual{}, and the manual's kills are a
strict subset of the LLM's, so within the status-oracle regime the LLM's \PDtestsLlm{}-test
volume reaches more of the catchable faults than the leaner \PDtestsManual{}-test suite. The
pilot's tie was the two levels coinciding through an authorship artefact.

\subsection{Implications}

\textbf{For practitioners.} An LLM generator gives breadth, not depth. It will reach every
documented operation and probe the documented error surface harder than a human or a search
tool. It will not notice that a pricing routine now returns the wrong
total, because nothing in the specification says what the right total is. The practical rule
is to use LLM generation for specification conformance and error-path breadth, and not to let
it displace regression oracles, property-based assertions or example-based expected values on
computational paths. EvoMaster found \RQIImcnemarLEc{} faults that the LLM missed entirely.

\textbf{For the field.} Every metric in our study that the surveyed literature actually uses
ranked the LLM first; the ground truth ranked it second. We doubt this is peculiar to our
subjects, since it is what scoring test suites on volume proxies produces. When AutoRestTest
reports 42 server errors against EvoMaster's 20~\cite{autoresttest}, or LlamaRestTest 204
faults against 130~\cite{llamaresttest}, those numbers are real, but they measure how hard the
documented error surface was hit, not what fraction of faults was found. Our results show that
the two can order arms oppositely on the same code. The per-endpoint scenario count that GAP-M
asks for, and that we introduce, carries the same caveat: RQ3 measures what the LLM writes,
and Section~\ref{sec:results-pvd} shows how little that predicted about what it caught.

\textbf{For benchmark design.} The field needs pre-seeded-fault benchmarks because without a
denominator the two orderings above are indistinguishable. We would add that a benchmark
should stratify its fault population by oracle-observability, separating status-visible from
value-only faults, since our results show that single axis reordering the arms. Otherwise the
fault mix decides the winner: a benchmark of only value faults will favour regression-oracle
tools, and one of only status faults will favour LLMs.

\subsection{Why the Low Absolute Recall Is Informative}

All three arms score low in absolute terms. The best, EvoMaster, kills \PDkilledEvo{} of
\PDkilledTotal{} mutants (\RQIIrecallEvoPct\%). We could have tuned this figure upward by
seeding easier faults, and chose not to, because the low ceiling is itself a finding about
black-box REST testing. Our mutants live in core logic reachable only through the HTTP
contract, and many are
equivalent at the boundary: the value they corrupt never surfaces in a response field, or
surfaces only for inputs that no arm generated. A white-box tool with code-coverage feedback
would score higher. Black-box REST suites, whether written by an LLM, a human or a search
algorithm, detect a small minority of seeded logic faults, and the interesting question is
which fault classes all three systematically cannot see.
