\section{Threats to Validity}
\label{sec:threats}

\subsection{Construct Validity}

\textbf{The manual arm is independent-agent, not independent-human.} Our proposal specified an
independent human baseline. No human cohort was available, so the manual suite was re-authored
by a separate, isolated model session with no shared context with the LLM suite, using
spec-only input and blind to both the LLM output and the fault catalog. This resolves the
pilot's same-author confound, in which one session wrote both suites and they killed identical
mutant sets. With the suites separated, the LLM-versus-manual comparison becomes informative
and the arms diverge ($b=\RQIImcnemarLMb$, $c=\RQIImcnemarLMc$;
Section~\ref{sec:results-rq2}). What remains uncontrolled is the gap between an independent
agent and an independent human tester. A human cohort might bring domain intuition that an
EP/BVA-from-spec agent lacks, and could plausibly close some of the LLM's lead. We therefore
state the RQ2 manual result as ``LLM versus an independently authored, specification-derived
EP/BVA suite'', not ``LLM versus human testers''. The direction of the conclusion is
unaffected, since the LLM's kills are a superset of the manual suite's.

\textbf{The LLM-versus-EvoMaster significance hinges on one boundary mutant.}
Our re-execution reproduces the pilot's EvoMaster kills exactly (the same 12
mutants on \texttt{rest-ncs}) and the LLM's \texttt{rest-scs} kills exactly, but on
\texttt{rest-ncs} the LLM's Recall falls from the pilot's 1/70 to \RQIINcsLlm{}. One LLM test
exercising \texttt{bessj} at the boundary $n=2$ passed on the pilot's build of the original SUT
but fails on our clean rebuild, so it is excluded as an invalid oracle and no longer kills that
mutant. Bessel-function evaluation at the boundary is numerically sensitive, and we regard
this as a genuine reproduction limit and not a pipeline error, since the exact EvoMaster
match rules out a build fault. Losing that one kill shifts the LLM-versus-EvoMaster discordant
count from five to $\RQIImcnemarLEb$, which is what carries the McNemar $p$ across the
$\alpha=\alphaRaw$ threshold, from just above it in the pilot to $\RQIImcnemarLEp$ here. The
direction (EvoMaster $>$ LLM) is robust either way. The $\alpha=\alphaRaw$ verdict is not, and
should be read as ``EvoMaster leads, at the edge of significance'' rather than as a firm
rejection.

\textbf{Oracle strength differs across arms by construction.} EvoMaster's recorded-value
regression oracles are strictly stronger on value faults than the spec-derived status oracles
of the LLM and manual arms. We did not fail to control this; it is the independent variable
that Section~\ref{sec:discussion} identifies as driving the result. It does mean, however, that
RQ2 does not compare generation techniques with oracle strength held constant. A fairer
isolation of generation quality would fix the oracle kind across arms and vary only the
generator. We report kill-by-subject so that the interaction stays visible instead of being
buried in a pooled average.

\textbf{The error-behaviour denominator is arm-derived, so its ceiling is not informative.}
The \PDerrKey{} HTTP error behaviours of Section~\ref{sec:results-pvd} are the union of what
the three arms elicited, so the highest-scoring arm reaches \PDerrKey{}$/$\PDerrKey{} by
construction. The LLM's score on this metric establishes that its observed behaviour set
contains the manual suite's and EvoMaster's, and nothing stronger. In particular it is not
evidence that the LLM covered the error surface, and it is not comparable to the code-derived
static baseline, which counts source locations rather than observed behaviours. A study
wanting a true error-surface Recall would have to score every arm against the static
enumeration; we did not, and we flag it as the clearest measurement gap in this paper.

\textbf{Mutation kills are a proxy for real faults.} A seeded single-token mutation is not a
field defect, and coverage is likewise a proxy for test quality. We report Recall, coverage
and edge-case counts jointly, and Section~\ref{sec:results-pvd} exists because any one of them
alone gives the wrong ranking.

\subsection{Internal Validity}

\textbf{Equivalent mutants are not excluded.} We do not attempt to identify semantically
equivalent mutants, which is undecidable in general. Some fraction of the \RQIIn{} mutants may
be unkillable by any black-box suite, which deflates all three arms' Recall uniformly. This
weakens the absolute figures but not the comparison, since every arm faces the identical
population.

\textbf{LLM non-determinism.} Generation used temperature 0 and a fixed seed, and raw outputs
are frozen in the repository, but LLM sampling is not bit-reproducible across provider-side
model updates. The generated suites are committed, so the analysis is reproducible even where
regeneration is not.

\textbf{Mutation confined to controller and core-logic classes.} We seed where faults are
observable through the HTTP contract. Faults in configuration, persistence or serialisation
layers fall outside the population and outside our conclusions.

\subsection{External Validity}

\textbf{Three JVM APIs, \RQIopsTotal{} operations.} Our subjects are small EMB services, so the
findings are EMB-only and JVM-only and should not be read as covering large production APIs,
authenticated flows or non-JVM stacks. We excluded the \texttt{user-management} service
(MySQL-backed) on setup cost, leaving authentication endpoints under-represented, so RQ1's
miss profile cannot speak to the endpoint type most likely to be missed.

\textbf{The \texttt{features-service} imbalance.} That subject contributes 4 mutants against 70
and 59 for the others, because its controllers contain little arithmetic to mutate. All three
arms tie there (\RQIIFeaturesLlm{} each), so it contributes almost nothing to the pooled
comparison. We keep it in, because dropping a subject after seeing its results would amount
to selecting on the outcome.

\textbf{One LLM, one configuration.} We used Claude Sonnet 4.6 at temperature 0 with a single
frozen prompt, so our results do not generalise to the GPT family that dominates the literature
(Section~\ref{sec:related}), to open-source models, or to agentic multi-turn strategies. Since
the mechanism we identify is oracle kind, we expect the qualitative finding to hold for any
spec-only generator. We have not tested that.

\subsection{Conclusion Validity}

\textbf{The Friedman test is underpowered by design.} With $n=3$ subjects it cannot detect
realistic effects ($\chi^2=\RQIIfriedmanChi$, $p=\RQIIfriedmanP$), which is why the proposal
pre-designated the pooled per-fault McNemar test ($N=\RQIIn$) as the primary RQ2 analysis. We
did not promote it after the omnibus test failed.

\textbf{RQ2's pairwise tests are significant but fragile.} Both McNemar comparisons reject at
$\alpha=\alphaRaw$ ($p=\RQIImcnemarLEp$ for LLM-versus-EvoMaster and $p=\RQIImcnemarLMp$ for
LLM-versus-manual), but neither clears the Bonferroni threshold $\alphaBonf$, and approximate
achieved power is below conventional adequacy for both: \RQIImcnemarLEpower{} for the
EvoMaster comparison and \RQIImcnemarLMpower{} for the manual one. We therefore rest the RQ2 conclusions on direction and effect size, namely that
EvoMaster kills more than twice as many mutants as the LLM and dominates outright on the
value-fault subject, and that the LLM's kills strictly contain the manual suite's. The
EvoMaster comparison is not a set containment: $b=\RQIImcnemarLEb$ mutants are killed by the
LLM alone. We do not rest the conclusions on the $p$-values alone, and we do not upgrade
``significant at $\alphaRaw$'' to ``firmly established''. The power figure is itself an approximation, since
statsmodels has no exact McNemar power routine and we substitute a two-independent-proportion
calculation on the same proportions, disclosed in \texttt{summary.json} alongside the value.

\textbf{Multiplicity.} Three RQ families are tested. We apply Holm correction within RQ2's
post-hoc comparisons and report a Bonferroni threshold across families
($\alpha_{\text{adj}}\approx\alphaBonf$). RQ1 and RQ3 survive both comfortably
($p=\RQIpvalue$ and $p=\RQIIIpvalue$). The RQ2 pairwise tests survive the raw threshold but not
Bonferroni, which is why we read them as directional evidence and not as decisive
rejections.
