\section{Conclusion}
\label{sec:conclusion}

We ran the comparison the REST-API-testing literature has not: an LLM generator, a manual
EP/BVA suite, and EvoMaster, on the same three APIs, against a \emph{known} population of
\RQIIn{} seeded faults rather than a tally of errors observed on a live system.

The answer to ``can an LLM out-test a human baseline and a search-based tool'' is: it
out-tests the human-style baseline and is out-tested by the tool. On every metric the field
currently reports, the LLM leads both: \RQIcoveragePct\% endpoint coverage ($p=\RQIpvalue$),
\RQIIItotalLlm{} edge-case scenarios against \RQIIItotalManual{} ($p=\RQIIIpvalue$,
rank-biserial $\RQIIIrankBiserial$), \PDtestsLlm{} test cases to EvoMaster's \PDtestsEvo{},
and \PDerrTriggeredLlm{} of \PDerrKey{} documented HTTP error behaviours triggered. On the
ground truth the ordering is EvoMaster \RQIIrecallEvo{} $>$ LLM \RQIIrecallLlm{} $>$ manual
\RQIIrecallManual{}: the LLM significantly out-detects an independently authored EP/BVA suite
($p=\RQIImcnemarLMp$, its kills a strict superset), yet is itself out-detected by EvoMaster
($p=\RQIImcnemarLEp$; both at the raw threshold, neither under Bonferroni, so directional
rather than decisive).

The explanation is not about model capability. A generator that reads a specification writes
assertions about \emph{status codes}, and a fault that changes a computed value while leaving
the status intact is invisible to every such assertion, no matter how many are written.
EvoMaster's recorded-value regression oracles see exactly those faults. The evidence is that
the arms' ordering tracks fault character rather than generator sophistication: on the
arithmetic-heavy \texttt{rest-ncs} \emph{both} status-oracle arms score \RQIINcsLlm{} while
EvoMaster kills \RQIINcsEvo{}; on the validation-heavy \texttt{rest-scs} the LLM leads every
arm. Two levels compose: oracle \emph{kind} fixes the ceiling (status oracles cannot reach
EvoMaster on value faults), and within a kind, \emph{volume} decides how much of the
reachable fault set is caught --- which is why the LLM's \PDtestsLlm{} tests beat the leaner
independent manual's \PDtestsManual{} where the faults are status-visible.

Three consequences follow. For practitioners, LLM generation is a breadth instrument that
complements rather than replaces value-based oracles --- it beats a human-style suite yet
still missed more than half the faults EvoMaster's regression oracles caught. For the field,
the metric that ranks the LLM first over EvoMaster (volume) and the metric that ranks it
below (ground-truth Recall) disagree on the same code, so the comparative claims currently in
the literature are not measuring what their framing implies. For benchmark design, fault
populations should be stratified by oracle-observability, because that single axis reorders
the arms.

Our principal remaining limitation is that the independent manual baseline is an isolated
\emph{agent} session, not a human cohort; a human replication against the same \RQIIn{}
mutants is the highest-value next step, and our pipeline runs it without modification. Beyond
it, the natural studies are an oracle-controlled comparison that fixes assertion kind and
varies only the generator, a fault population stratified by oracle-observability, and the
GAP-T ablation across model families. The pipeline, prompts, fault catalogs, and raw data are
committed, and every number above regenerates from them by script.
