% =====================================================================
% ACTION REQUIRED BEFORE SUBMISSION
% This section states, on the authors' behalf, how AI tools were used.
% Springer requires that such use be disclosed and that LLMs are NOT
% listed as authors. The text below describes the uses visible in the
% repository. VERIFY EVERY SENTENCE against what the team actually did
% and edit it to match; an inaccurate disclosure is worse than none.
% =====================================================================

\section*{Use of AI Tools}
\label{sec:ai-disclosure}

This study concerns LLM-based test generation, so we separate three distinct roles that
language models play in it.

\textbf{As an object of study.} Claude Sonnet 4.6 is the subject of the LLM arm. Its
configuration, prompt and outputs are specified in Section~\ref{sec:method} and committed to
the replication package.

\textbf{As an instrument in the experiment.} The manual EP/BVA arm was authored by a separate,
isolated model session rather than by a human cohort, because no human cohort was available.
We treat this as a construct-validity threat, label the arm ``independently authored'' rather
than ``human'' throughout, and discuss its consequences in Section~\ref{sec:threats}.

\textbf{In manuscript preparation.} The authors used an LLM assistant to draft and edit prose,
and to convert the manuscript to this template. All research design, experimental execution,
data analysis and interpretation were carried out by the authors. Every computed statistic
reported here is generated from committed raw data by \texttt{scripts/gen\_paper\_macros.py}
and was not written by a model or transcribed by hand (Section~\ref{sec:method-repro}). The
authors read,
verified and take full responsibility for the entire content of this paper. No LLM is listed
as an author, in line with Springer's authorship policy.

\section*{Data Availability}

The replication package contains the mutation catalogs, the generated test suites for all
three arms, the raw kill matrices, the analysis scripts and the notebook that regenerates
every statistic in this paper. Subjects are drawn from the EvoMaster Benchmark (EMB,
LGPL-3.0); the upstream commit used is recorded in the repository.

% LNCS requires Disclosure of Interests in the `credits' environment, using the
% \discintname macro for the run-in heading (llncsdoc v2.26 sec. 5). It is mandatory,
% and must sit immediately before the bibliography.
\begin{credits}
\subsubsection{\discintname}
The authors have no competing interests to declare that are relevant to the content of this
article.
\end{credits}
